Measuring matter effects in Neutron star - Black hole binaries with Advanced LIGO

### 2016-01-29 Paper/paper.tex
added new figures. IMP

\begin{abstract}
We attempt to answer the following questions: 
\begin{enumerate}
 \item Will ignoring the effect of NS deformability during inspiral and NS 
 disruption during merger (in an NSBH binary) change the efficiency of our 
 searches?
 \item What systematic biases are introduced by neglecting the NS tidal
 disruption close to merger? 
 \item At what SNR do the systematic biases dominate over the statistical 
 errors in the recovery of binary parameters?
 \item By using templates which models NS deformation (over inspiral) and 
 NS disruption (close to merger), what information can we recover about
 the $\Lambda_\mathrm{BH}\sim (R/M)^5_\mathrm{NS}$ parameter?
\end{enumerate}
\end{abstract}

@@

NS-BH binaries are one of the most important sources of gravitational waves (GW)
with these ground-based instruments, and we expect to detect $\mathcal{O}(10)$
of such systems every year~\cite{Abadie:2010cf}. 
%
One question that aLIGO will likely probe is `what is the nature of
matter at nuclear densities', i.e. what is the equation of state (EoS) of 
neutron stars. The EoS determines the compactness of NSs, which determines
their deformation over the course of inspiral (albeit weakly, at $5$PN order in 
phasing~\prayush{CITE PN TIDAL RESULTS}), as well as how early in the coalescence
process will the NS be disrupted by the tidal field of the BH. Once disrupted,
the NS forms an accretion disk around the BH. Most of this matter disk spirals 
into the BH within a few milliseconds, with a small fraction becoming unbounded
and gets ejected~\prayush{CITE BHNS NR}. It is believed that an accompaning
short (lasting less than $2$ seconds) Gamma-ray Burst (sGRB) might be present
under certain physical conditions~\prayush{CITE NSBH-GRB},
making these systems one of the prime targets for joint electromagnetic-GW
astronomy searches~\prayush{CITE EM-GW}.
In such a situtation, the emitted GW signal
is heavily damped post-disruption, and the quasi-normal modes of the resulting
BH are not excited. Such a GW signal is qualitatively very different from a
BH-BH merger, and one of the main goals of this work is to determine the 
signal strength at which this difference will be quantifiabe by aLIGO.
%
However, studies have shown that for mass-ratio $\mbh/\mns\gtrsim 5$ the BH
will need to have fairly large and positively-aligned (with the orbit) spins
in order for its tidal field to be strong enough to disrupt the NS before the
innermost stable circular orbit (ISCO) is reached~\cite{Foucart:2013psa}.

Conventional astronomical 
methods have, to date, made observations of NSs with masses between 
$XX-YY M_\odot$ and spins $|\vec{\chi}_\mathrm{NS}|\lesssim 0.05$. On the other hand, 
indirect observations of stellar-mass BHs place their masses between 
$5-35M_\odot$, with their spin angular momenta $|\vec{\chi}_\mathrm{BH}|$ 
ranging from small to nearly extremal (Kerr) values (see,
e.g.~\cite{McClintockEtAl:2006,Miller:2009cw,Gou:2014una} for examples of 
nearly extremal estimates of BH spins, and see
Refs.~\cite{McClintock:2013vwa,Reynolds:2013qqa} for recent reviews of
astrophysical BH spin measurements).

@@

The goal of this paper is to probe the distinguishability of BHBH and NSBH
mergers, focusing at the effects of the BH mass, BH spin, and NS tidal 
deformability. We also investigate the accuracy with which aLIGO can measure
the tidal deformability parameter $\Lambda_\NS\sim(R/M)^5_\NS$. For both of 
these studies, we will use the zero-detuning high-power noise curve for the
aLIGO detectors. \prayush{In addition, we also study the effect of tuning aLIGO
over a narrow frequency range aimed at NSBH mergers.}
We use the models published in Ref.~\cite{Lackey:2013axa,Pannarale:2015jka} to
simulate NSBH merger waveforms.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{How are detection searches affected by not including NS matter effects}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Near future detection searches with aLIGO are poised to search for NSBH
signals using BHBH templates. In this section, we examine the effect of ignoring
the effects of tidal deformation and disruption of the NS during inspiral and 
merger phases of search templates. We construct a template bank $\mathcal{B}$ 
restricted to the NSBH parameter space with $\mbh\in[2.5M_\odot, 20.5M_\odot]$, 
$\chibh\in[-0.5, +0.75]$, $\mns\in[1M_\odot,3M_\odot]$ and $\chi_\mathrm{NS}=0$,
using a PN-based geometric lattice placement method~\cite{Harry:2013tca}.

Before investigating its SNR recovery against tidal NSBH signals, we test the 
effectualness of this bank to point-particle NSBH signals,
in order to make sure that the placement method does not leave any region of
the binary parameter-space under-covered. We simulate $50,000$ signals with 
component-mass and BH spin values
sampled uniformly over their allowed ranges, and filter each of them against
the entire bank. For both signals and templates, we use the frequency-domain 
reduced-order model (ROM) of the SEOBNRv2 approximant~\cite{Purrer:2014} 
implemented within the LIGO algorithms Library~\cite{lal}. We record the 
fitting-factors~\cite{FittingFactorApostolatos} recovered by the bank for all
signals, which measures the fraction of optimal SNR that the bank is able
to recover. Fig.~\ref{} shows these FFs as a function of the injected
BH mass and BH spin, and Fig.~\ref{} provides a histogram of the recovered FF 
values. From these figures, we conclude that our bank is effectual as intended.

Next, tn order to establish the effect of NS matter effects on the SNR recovery of 
detection searches, we perform a similar calculation as above, with the
difference that the signals are now modelled using the model of tidal 
corrections to IMR amplitude and phasing published in 
Ref.~\cite{Lackey:2013axa}, applied to SEOBNRv2 as the underlying BHBH model.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{What do we gain by using templates that include NS matter effects?}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % 
% \begin{figure*}
% \centering
% \textbf{NS-BH Injection; NS-BH Templates}
% \includegraphics[width=1.7\columnwidth]{plots/TT_NSLambdaBias_vs_SNR_q23.pdf}
% \caption{This figure is similar to the bottom row of
% Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23}, with the difference that  we
% consider the NS tidal parameter $\Lambda$ here.}
% \label{fig:TT_NSLambdaBias_vs_SNR_q23}
% \end{figure*}
% % 
% \begin{figure*}
% \centering
% \textbf{NS-BH Injection; NS-BH Templates}
% \includegraphics[width=1.7\columnwidth]{plots/TT_NSLambdaCIWidth90_vs_SNR_q23.pdf}
% \caption{This figure is similar to the bottom row of
% Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23}, with the difference that  
% we consider the NS tidal parameter $\Lambda$ here.}
% \label{fig:TT_NSLambdaCIWidth90_vs_SNR_q23}
% \end{figure*}
% % 
% \begin{figure*}
% \centering    
% \includegraphics[width=2\columnwidth]{plots-TT/TT_LambdaBiases_CI90_0_Lambda_SNR.png}
% \caption{This figure shows.}
% \label{fig:TT_LambdaBiases_CI90_0_Lambda_SNR}
% \end{figure*}

\begin{figure*}
\centering    
\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda500_0_CI68_3.png}
\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda800_0_CI68_3.png}
\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda1000_0_CI68_3.png}\\
\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda500_0_CI90_0.png}
\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda800_0_CI90_0.png}
\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda1000_0_CI90_0.png}
\caption{This figure shows.}
\label{fig:TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_CI90_0}
\end{figure*}

% 
\begin{figure*}
\centering    
\includegraphics[width=2.1\columnwidth]{plots-TT/TTLambdaCIWidths68_3_Lambda_SNR.png}\\
\includegraphics[width=2.1\columnwidth]{plots-TT/TTLambdaCIWidths90_0_Lambda_SNR.png}
\caption{This figure shows.}
\label{fig:TT_LambdaCIWidths90_0_Lambda_SNR}
\end{figure*}
% 
%\begin{figure*}
%\centering    
%\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda500_0_CI68_3.png}
%\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda800_0_CI68_3.png}
%\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda1000_0_CI68_3.png}\\
%\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda500_0_CI90_0.png}
%\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda800_0_CI90_0.png}
%\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda1000_0_CI90_0.png}
%\caption{This figure shows.}
%\label{fig:TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_CI90_0}
%\end{figure*}
%% 
%\begin{figure*}
%\centering    
%\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR30_0_CI68_3.png}
%\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR50_0_CI68_3.png}
%\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR90_0_CI68_3.png}\\
%\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR50_0_CI90_0.png}
%\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR90_0_CI90_0.png}
%\caption{This figure shows.}
%\label{fig:TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass}
%\end{figure*}

@@

\section{How is PE affected by not including NS matter effects?}
% \begin{figure*}
% \centering 
% % \textbf{BH-BH Injection; BH-BH Templates}
% % \includegraphics[width=1.7\columnwidth]{plots/NN_chirpMassBias_vs_SNR_q23.pdf}\\ 
% \textbf{NS-BH Injection; BH-BH Templates}
% \includegraphics[width=1.7\columnwidth]{plots/TN_chirpMassBias_vs_SNR_q23.pdf}\\ 
% \textbf{NS-BH Injection; NS-BH Templates}
% \includegraphics[width=1.7\columnwidth]{plots/TT_chirpMassBias_vs_SNR_q23.pdf}%\quad
% \caption{These figures show the (fractional) difference between the chirp mass 
% value corresponding to the median of its posterior probability distribution,
% and the actual injected chirp mass,
% as a function of the signal-to-noise-ratio (SNR) of the injected signal.
% % In the top row, NS matter effects on the inspiral and merger waveform
% % are neithed included in templates nor in the injected signal. 
% In the top row, NS matter effects (inspiral and merger) are included in the
% injections but not in templates. This corresponds to the current 
% plan for aLIGO parameter estimation studies. The bottom row shows the effect
% of additionally including NS matter effects in the templates.
% % 
% In each row, the left and right panel correspond to 
% $q=m_\mathrm{BH}/m_\mathrm{NS}=\{2,3\}$, respectively. 
% In each panel, the color of each curve corresponds to the value of NS 
% deformability parameter $(\Lambda=\{500,1000,2000\})$, 
% and line-style corresponds to the value of
% the dimensionless spin on the BH ($\chi_\mathrm{BH}=\{-0.5,0,+0.5\}$).
% }
% \label{fig:TNT_chirpMassBias_vs_SNR_q23}
% \end{figure*}
% % 
% \begin{figure*}
% \centering 
% % \textbf{BH-BH Injection; BH-BH Templates}
% % \includegraphics[width=1.7\columnwidth]{plots/NN_chirpMassCIWidth90_vs_SNR_q23.pdf}\\ 
% \textbf{NS-BH Injection; BH-BH Templates}
% \includegraphics[width=1.7\columnwidth]{plots/TN_chirpMassCIWidth90_vs_SNR_q23.pdf}\\ 
% \textbf{NS-BH Injection; NS-BH Templates}
% \includegraphics[width=1.7\columnwidth]{plots/TT_chirpMassCIWidth90_vs_SNR_q23.pdf}%\quad
% \caption{These figures are similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23}
% with the difference that here we show the width of the $90\%$ confidence 
% interval (recovered) for chirp mass, as a function of the injected SNR. 
% Note that the confidence interval's width is normalized by the injected
% parameter value.
% }
% \label{fig:TNT_chirpMassCIWidth90_vs_SNR_q23}
% \end{figure*}
% % 
% \begin{figure*}[!t]
% \centering    
% \textbf{NS-BH Injection; BH-BH Templates}
% \includegraphics[width=1.7\columnwidth]{plots/TN_EtaBias_vs_SNR_q23.pdf}\\ 
% \textbf{NS-BH Injection; NS-BH Templates}
% \includegraphics[width=1.7\columnwidth]{plots/TT_EtaBias_vs_SNR_q23.pdf}\\%\quad
% \textbf{BH-BH Injection; BH-BH Templates}
% \includegraphics[width=1.7\columnwidth]{plots/NN_EtaBias_vs_SNR_q23.pdf} 
% \caption{This figure (top two rows) is similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
% with the difference that we consider the systematic bias in the dimensionless
% mass-ratio $\eta$ here. In the bottom row we show control runs where both
% signal and template waveforms lack matter effects.}
% \label{fig:TNT_EtaBias_vs_SNR_q23}
% \end{figure*}
% % 
% \begin{figure*}
% \centering    
% \textbf{NS-BH Injection; BH-BH Templates}
% \includegraphics[width=1.7\columnwidth]{plots/TN_EtaCIWidth90_vs_SNR_q23.pdf}\\ 
% \textbf{NS-BH Injection; NS-BH Templates}
% \includegraphics[width=1.7\columnwidth]{plots/TT_EtaCIWidth90_vs_SNR_q23.pdf}\\%\quad
% \textbf{BH-BH Injection; BH-BH Templates}
% \includegraphics[width=1.7\columnwidth]{plots/NN_EtaCIWidth90_vs_SNR_q23.pdf} 
% \caption{This figure (top two rows) is similar to Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
%   with the difference that we consider the confidence intervals in the recovery
%   of the dimensionless mass-ratio $\eta$ here. The bottom row here is similar to 

>>>>

\begin{abstract}
We attempt to answer the following questions: 
\begin{enumerate}
 \item What systematic biases are introduced by neglecting the NS tidal
 disruption close to merger? 
 \item At what SNR do the systematic biases dominate over the statistical 
 errors in the recovery of binary parameters?
 \item By using templates which models NS deformation (over inspiral) and 
 NS disruption (close to merger), what information can we recover about
 the $\Lambda_\mathrm{BH}\sim (R/M)^5_\mathrm{NS}$ parameter?
\end{enumerate}
\end{abstract}

@@

NS-BH binaries are one of the most important sources of gravitational waves (GW)
with these ground-based instruments, and we expect to detect $\mathcal{O}(10)$
of such systems every year~\cite{Abadie:2010cf}. 
%
One question that aLIGO will likely probe is `what is the nature of
matter at nuclear densities', i.e. what is the equation of state (EoS) of 
neutron stars. The EoS determines the compactness of NSs, which determines
their deformation over the course of inspiral (albeit weakly, at $5$PN order in 
phasing~\prayush{CITE PN TIDAL RESULTS}), as well as how early in the coalescence
process will the NS be disrupted by the tidal field of the BH~\prayush{[CITE]}.
Once disrupted,
the NS forms an accretion disk around the BH. Most of this matter disk spirals 
into the BH within a few milliseconds, with a small fraction becoming unbounded
and gets ejected~\prayush{CITE BHNS NR}. It is believed that an accompanying
short (lasting less than $2$ seconds) Gamma-ray Burst (sGRB) might be present
under certain physical conditions~\prayush{CITE NSBH-GRB},
making these systems one of the prime targets for joint electromagnetic-GW
astronomy searches~\prayush{CITE EM-GW}.
In such a situtation, the emitted GW signal
is heavily damped post-disruption, and the quasi-normal modes of the resulting
BH are not excited. Such a GW signal is qualitatively very different from a
BH-BH merger, and one of the main goals of this work is to determine the 
signal strength at which this difference will be quantifiabe by aLIGO.
%
However, studies have shown that for mass-ratio $\mbh/\mns\gtrsim 5$ the BH
will need to have fairly large and positively-aligned (with the orbit) spins
in order for its tidal field to be strong enough to disrupt the NS before the
innermost stable circular orbit (ISCO) is reached~\cite{Foucart:2013psa}.

Conventional astronomical methods have, to date, made observations of ~2500
NSs with masses between $1.25-2.1 M_\odot$~\cite{Lyne:2004cj,
Demorest:2010bx,2013Sci...340..448A,atnfcatalog,mcgillmagnetarcatalog,
stellarcollapsemass}, although
tightly clustered within $1.35\pm0.5M_\odot$~\cite{stellarcollapsemass}.
%
Observations of over 2500 NSs have shown that the spin of NSs
$|\vec{\chi}_\mathrm{NS}|$ have magnitudes that are $<1\%$~\cite{Miller:2014aaa}.
%
On the other hand, indirect observations of stellar-mass BHs place their
masses between $5-35M_\odot$, with their spin angular momenta 
$|\vec{\chi}_\mathrm{BH}|$ ranging from small to nearly extremal (Kerr) values
(see, e.g., Refs.~\cite{McClintockEtAl:2006,Miller:2009cw,Gou:2014una} for 
examples of nearly extremal estimates of BH spins, Refs.~\cite{McClintock:2013vwa,
Reynolds:2013qqa} for recent reviews of astrophysical BH spin measurements,
and Figure 5 of Ref.~\cite{Miller:2014aaa} for a comparison of NS and BH spins).

@@

The goal of this paper is to probe the distinguishability of BHBH and NSBH
mergers, focusing at the effects of the BH mass, BH spin, and the tidal 
deformability parameter $\Lambda_\NS\sim(R/M)^5_\NS$. We also investigate the
accuracy with which aLIGO can measure $\Lambda_\NS$ itself. For both of 
these studies, we will use the zero-detuning high-power noise curve for the
aLIGO detectors. \prayush{In addition, we also study the effect of tuning aLIGO
over a narrow frequency range aimed at NSBH mergers.}
We use the models published in Ref.~\cite{Lackey:2013axa,Pannarale:2015jka} to
simulate NSBH merger waveforms.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{How are detection searches affected by not including NS matter effects}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%Near future detection searches with aLIGO are poised to search for NSBH
%signals using BHBH templates. In this section, we examine the effect of ignoring
%the effects of tidal deformation and disruption of the NS during inspiral and 
%merger phases of search templates. We construct a template bank $\mathcal{B}$ 
%restricted to the NSBH parameter space with $\mbh\in[2.5M_\odot, 20.5M_\odot]$, 
%$\chibh\in[-0.5, +0.75]$, $\mns\in[1M_\odot,3M_\odot]$ and $\chi_\mathrm{NS}=0$,
%using a PN-based geometric lattice placement method~\cite{Harry:2013tca}. 
%
%
%Before investigating its SNR recovery against tidal NSBH signals, we test the 
%effectualness of this bank to point-particle NSBH signals,
%in order to make sure that the placement method does not leave any region of
%the binary parameter-space under-covered. We simulate $50,000$ signals with 
%component-mass and BH spin values
%sampled uniformly over their allowed ranges, and filter each of them against
%the entire bank. For both signals and templates, we use the frequency-domain 
%reduced-order model (ROM) of the SEOBNRv2 approximant~\cite{Purrer:2014} 
%implemented within the LIGO algorithms Library~\cite{lal}. We record the 
%fitting-factors~\cite{FittingFactorApostolatos} recovered by the bank for all
%signals, which measures the fraction of optimal SNR that the bank is able
%to recover. Fig.~\ref{} shows these FFs as a function of the injected
%BH mass and BH spin, and Fig.~\ref{} provides a histogram of the recovered FF 
%values. From these figures, we conclude that our bank is effectual as intended.
%
%
%Next, tn order to establish the effect of NS matter effects on the SNR recovery of 
%detection searches, we perform a similar calculation as above, with the
%difference that the signals are now modelled using the model of tidal 
%corrections to IMR amplitude and phasing published in 
%Ref.~\cite{Lackey:2013axa}, applied to SEOBNRv2 as the underlying BHBH model.
\section{What do we gain by using templates that include NS matter effects?}
%\begin{figure*}
%\centering    
%\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda500_0_CI68_3.png}
%\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda800_0_CI68_3.png}
%\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda1000_0_CI68_3.png}\\
%\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda500_0_CI90_0.png}
%\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda800_0_CI90_0.png}
%\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda1000_0_CI90_0.png}
%\caption{This figure shows.}
%\label{fig:TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_CI90_0}
%\end{figure*}
\begin{figure*}
\centering    
\includegraphics[trim={3cm 0 0 0},width=2.3\columnwidth]{plots/TTLambdaCIWidths90_0_Lambda_SNR.pdf}
%\includegraphics[trim={3cm 0 0 0},width=2.3\columnwidth]{plots/TTLambdaCIWidths90_0_Lambda_SNR.pdf}
\caption{This figure shows the statistical uncertainty in the measurement of
$\lambdans$, as a percentage of the injected signal's $\lambdans$. In each panel, 
the same is shown as a function of the BH mass and spin, keeping the 
$\lambdans$ and injection's SNR $\rho$ fixed (given in the panel). Each row contains
panels with the same value of $\lambdans$, while each column contains panels 
with the same value of $\rho$. Two contours are drawn in each panel demarkating
that the statistical measurement uncertainty for $\lambda$ is equal to (or twice
of) its injected $\lambdans$ value, i.e. where we can place a factor of $1-2$ 
bound on $\lambdans$.
We note that, as expected, the measurement accuracy for $\lambdans$ improves with (i) increasing
SNR, (ii) decreasing BH mass, (iii) increasing BH spin, and 
(iv) increasing $\lambdans$, i.e. the tidal deformability of the neutron star.
}
\label{fig:TT_LambdaCIWidths90_0_Lambda_SNR}
\end{figure*}
\begin{figure*}
\centering    
\includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR20_CI90_0.pdf}
\includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR30_CI90_0.pdf}
\includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR50_CI90_0.pdf}\\
% \includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR70_CI90_0.pdf}
% \includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR90_CI90_0.pdf}
% \includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR120_CI90_0.pdf}
\caption{We demarkate here the regions of BH mass and spin plane where we can 
constrain the $\lambdans$ of its companion NS within a factor of $1-2$ of its true value.
Panels correspond to the injected SNR, with signal strength increasing
from left to right and top to bottom. Within each panel, $\lambdans$ measurement uncertainty
contours are shown as functions of BH mass and spin, with different colors corresponding to the
injected $\lambdans$. We note that this information is, in principle, contained in 
Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, which we gather here to better understand
the effect of NS's deformability itself on its measurement. As from Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, we note that the measurability of NS matter
effects improves with all factors that enhances the signature of the NS's disruption on
the \textit{detectable} portion of the emitted GW signal (implying, within a frequency band
set by the detectors).
}
\label{fig:TT_LambdaErrorCurves_BHspin_BHmass_CI90_0}
\end{figure*}
\begin{figure*}
\centering    
\includegraphics[trim={0 0 0 0},width=\columnwidth]{plots/TT_q2-00_mNS1-35_chiBH0-75_Lambda1500-0_SNR20-0_corner_eta_Mc_chi2_Lambda.pdf}
\includegraphics[trim={0 0 0 0},width=\columnwidth]{plots/TT_q2-00_mNS1-35_chiBH0-75_Lambda1500-0_SNR20-0_corner_m1_m2_chi2_Lambda.pdf}
\caption{These figure shows the $1-$ and $2-$dimensional 
\textit{a posteriori} probability distributions 
for different physical parameters describing a BHNS system including NS matter 
effects, as measured by a Bayesian Markov-chain Monte Carlo using tidal templates.
The true parameters of the injected signal are: 
$\mathcal{M}=1.6406M_\odot, \eta=0.2222, \chibh=\chi_2=0.75$ and $\lambdans=1500$,
with the injected SNR $\rho=20$. We show this case as typical of those with the largest
tidal signatures imprinted on the emitted GWs (due to the system being comparable
mass-ratio with a large-aligned spin on the BH and a fairly deformable NS), yet 
with low SNR.
We note a few things:
(i) chirp mass is well constrained, with a nearly-Gaussian probability distribution;
while (ii) the same for the other $3$ less constrained parameters rails against
prior boundaries. The specific boudary conditions that we run into are (a) for $\eta$,
we are limited by the requirement that $m_\mathrm{NS}\geq 1.2M_\odot$ 
of our templates; (b) for BH spin, we are limited by the $\chibh\leq+0.75$ requirement
that results from a template model limitation~\cite{Lackey:2013axa}; and (c) for 
$\lambdans$, we
restrict it to $[0, 4000]$ for the same reason as (b). These prior boundaries
limit the exploration of the entire likelihood surfaces. While (a) is astrophysically
justified, given that we have not observed NSs with masses below 
$1.25M_\odot$, (b) and (c) will be removed in future studies as more evolved template
models become available. See, also, Fig.~\ref{fig:TT_logLikeProb_q2_chi75_L1500_SNR20}.
Finally, we note that the left panel corresponds to the actual sampling parameters,
while the right panel to derived ones.
}
\label{fig:TT_pdf_q2_chi75_L1500_SNR20}
\end{figure*}
\begin{figure*}
\centering    
% \includegraphics[width=.66\columnwidth]{plots/TT_q2-00_mNS1-35_chiBH0-75_Lambda1500-0_SNR20-0_corner_eta_Mc_chi2_Lambda.pdf}
\includegraphics[width=\columnwidth]{plots/TT_logLike_vs_m1_m2_eta0-22_Mc1-64_chi0-00_chi0-50_L1000-0_SNR30.pdf}
\includegraphics[width=\columnwidth]{plots/TT_logProb_vs_m1_m2_eta0-22_Mc1-64_chi0-00_chi0-50_L1000-0_SNR30.pdf}
\caption{These figures show the effect of limitation (a) (described in the caption
of Fig.~\ref{fig:TT_pdf_q2_chi75_L1500_SNR20}) on the likelihood surface portion
that is actually accessible to our Bayesian inferencing algorithm. On the left we show the
entire likelihood surface in the $\mbh-\mns$ plane, while on the left is the 
portion not forbidden by the requirement $\mns\geq 1.2M_\odot$. Similarly, our
priors on BH spin and NS deformability further curtail the portion of sampled
parameter space, albeit to much less impact than this mass cut.}
\label{fig:TT_logLikeProb_q2_chi75_L1500_SNR20}
\end{figure*}

%
%% 
%\begin{figure*}
%\centering    
%\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR30_0_CI68_3.png}
%\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR50_0_CI68_3.png}
%\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR90_0_CI68_3.png}\\
%\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR50_0_CI90_0.png}
%\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR90_0_CI90_0.png}
%\caption{This figure shows.}
%\label{fig:TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass}
%\end{figure*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{How is PE affected by not including NS matter effects?}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% #################
\begin{figure*}
\centering 
%\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI68_3_Lambda_SNR.pdf}\\
\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR}.pdf
\caption{We show here (Log of) the ratio of the systematic and statistical
measurement uncertainties for the binary chirp mass over the NSBH parameter 
space. Each panel shows the same as a function of BH mass and spin. Across
each row, we can see the effect of increasing the signal strength (i.e. SNR)
with the NS's deformability fixed. Down each column, we can see 
the effect of the increasing NS's tidal deformability at fixed SNR. We also
show dashed contours for where the ratio equals $10\%$ (labelled ``-1'')
or $100\%$ (labelled ``0'').
% 
Generally, these two sources of uncertainty are different and for BBHs, the
statistical errors dominate systematic ones~\cite{Kumar:2016dhh}. We find
that for BHNS binaries, its not much different until we get to high SNRs
$\rho\gtrsim 70$ for which if the BH is spinning rapidly enough, the 
measurement uncertainty for $\mchirp$ could become dominated by the exclusion
of tidal effects in filtering templates.
}
\label{fig:TN_chirpMassBias_vs_Lambda_SNR}
\end{figure*}
%
\begin{figure*}[!t]
\centering    
%\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI68_3_Lambda_SNR.pdf}\\
\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR.pdf}\\
\caption{These figures are similar to Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}
with the difference that the quantity in question here is the symmetric 
mass-ratio $\eta$. We find that for fairly loud GW signals, with $\rho\simeq 50$,
NS tidal effects on GW emission can become the dominant source of error. However,
for quieter signals with $\rho\leq 30$, we find that not including NS matter 
effects in parameter estimation templates has minimal effects on the measurement
of $\eta$. We remind the reader that the SNRs here are all single detector values.
}
\label{fig:TN_EtaBias_vs_Lambda_SNR}
\end{figure*}
% 
% 
\begin{figure*}
\centering
\includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI68_3.pdf}
\includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI90_0.pdf}
\caption{These figures show the ratio of the systematic and statistical
measurement errors for BH spins, for all of our injected cases. The left
panel corresponds to $1-\sigma$ statistical confidence level, while the right panel
corresponds to $90\%$ confidence level. We note that below $\rho\approx 30$, ignoring
tidal effects in templates introduces systematic effects that remain subdominant
to the statistical measurement uncertainties.
}
\label{fig:TNT_BHspinBias_vs_SNR_q23}
\end{figure*}
% 
% 
Advanced LIGO searches and parameter estimation efforts aimed at binaries 
containing NS and stellar-mass BHs are poised to use waveform models that do not
include the effects of the NS tidal deformability~\cite{Canton:2014ena}.
While this is not expected to
be the dominant source of error at low SNRs, it will likely introduce a
systematic bias in the physical parameters that we recover from GW observations.
In this section, we study the impact of the same for binaries where the BH spins
are aligned to the orbital angular momentum.
